Lies and the lying liars who tell them

### 2018-11-01 GSV-comment-brief.Rmd
Make LaTeX (ugh) work

---
title: "Lies and the lying liars who tell them: comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
output:
  bookdown::tufte_book2:
    citation_package: natbib
    latex_engine: xelatex
    md_extensions: +raw_attribute
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
bibliography: bibliography.bib
link-citations: yes
---

@@

library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE)
options(huxtable.knitr_output_format = "html")
source("bayesian-heads.R")
suppressPackageStartupMessages({
  library(Hmisc)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(broom)
  library(huxtable)
})

huxtable::set_default_properties(position = "left")
set.seed(19751027)

@@

I ran simulations to check the overall performance of the GSV confidence intervals. Simulations
parameters are shown in Table \@ref(tab:params-table). For each parameter combination, I ran 1000
simulations. "Liars" always reported heads. Non-liars reported tails with probability $P$ and
heads otherwise.

@@

huxtable(
        Parameter = c(
          "Sample size (N)", 
          "Proportion of liars (L/N)", 
          "Probability of bad random outcome (P)"
        ),
        Values    = c(
          paste(params$N, collapse = ", "),
          paste(params$Liars, collapse = ", "),
          paste(params$P, collapse = ", ")
        ),
        add_colnames = TRUE
      ) %>% 
      set_caption("(#tab:params-table) Parameter values") %>% 
      theme_article()
```

For each simulation and confidence interval, I computed whether the confidence interval 
contained the true value of $L/N$. Table \@ref(tab:coverage-overall) shows the results.

@@

cov_overall <- coverage %>% 
      ungroup() %>% 
      mutate(
        CI = paste0("CI ", CI, "%"),
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      spread(CI, mean_in_ci) %>% 
      arrange(factor(method, levels = c("GSV", "Frequentist", "Bayesian")))

@@

ht_overall <- as_huxtable(cov_overall, add_colnames = TRUE) %>% 
      set_caption("(#tab:coverage-overall) Proportion of true results within confidence interval.") %>% 
      set_label("tab:coverage-overall") %>% 
      set_number_format(-1, -1, list(pct_format)) %>% 
      theme_article()

@@

The problem is especially bad when there are many reports of heads. In this case there were probably
many true heads, so $T$ is high and the true sample size $N-T$ is low. Table \@ref(tab:coverage-rn)
shows this. It splits the simulations by the proportion of reported heads, $R/N$. Coverage levels
fall off sharply as $R/N$ increases. Note that for fair coin flips, $R/N$ is typically greater than
0.5, in the simulations and in reality.

```{r coverage-rn}
coverage_rn <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = factor(Hmisc::cut2(R/N, cuts = c(0, 1/4, 1/2, 3/4, 1)))      
      ) %>% 
      mutate(pct_cases = n()/nrow(.)) %>% 
      group_by(`R/N:`, CI) %>% 
      summarize(
        mean_in_ci  = mean(within_ci, na.rm = TRUE), 
        `Pct cases` = pct_cases[1]
      ) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci)

as_huxtable(coverage_rn, add_colnames = TRUE) %>% 
      set_caption("(#tab:coverage-rn) GSV confidence interval coverage by low and high R/N") %>% 
      set_label("tab:coverage-rn") %>% 
      set_escape_contents(2:3, 1, FALSE) %>% 
      set_number_format(-1, -1, list(pct_format)) %>% 
      theme_article()
```

@@

Results are shown in Table \@ref(tab:coverage-allmethods). Both frequentist and  Bayesian methods
mostly achieve the nominal confidence level, with more than 90/95/99% of intervals containing the
true value of $Liars$. The exception is the frequentist 99% confidence interval, which is too
narrow.

```{r coverage-allmethods}
ht_overall %>% 
      set_caption("(#tab:coverage-allmethods) Coverage levels for GSV and alternative methods")
```

Frequentist confidence intervals could be less accurate when N is low, since that leads to more
sampling variation in the number of true heads. Table \@ref(tab:coverage-N) checks this by looking
separately at simulations with $N=10$ and $N=50$. Frequentist 99% confidence intervals indeed appear
slightly too narrow for this range. Bayesian confidence intervals are fine. Both alternatives
perform much better than the GSV approach.

@@

as_huxtable(coverage_N, add_colnames = TRUE) %>% 
      set_rowspan(c(2, 4, 6), "Method", 2) %>% 
      set_number_format(-1, contains("CI"), list(pct_format)) %>% 
      set_caption("(#tab:coverage-N) Confidence interval coverage by sample size") %>% 
      set_label("table:coverage-N") %>% 
      theme_article()
```

@@

We can also compare the accuracy of point estimates of $\textit{Liars}$. Table
\@ref(tab:point-estimates) shows the mean squared error for methods by different $N$. For low $N$,
the best method is Bayesian and the worst is Frequentist, with GSV in between. When $N$ gets large all
methods give about the same estimates and are equally accurate.

@@

sim_est %>% 
      filter(CI == 90) %>% 
      group_by(method, N) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("(#tab:point-estimates) Mean squared errors by method and N") %>% 
      theme_article()
```

@@

```{r hj2016-pic, fig.cap = "Posterior means and confidence intervals for proportion dishonest in 15 countries. Crosses are naive estimates."}

@@

Results are shown in Figure \@ref(fig:hj2016-pic), with means and 95% confidence intervals. There
is substantial shrinkage from the "naive" per-country estimates (using $2R/N - 1$). One of the strengths of empirical Bayes,  as @casella1985introduction points out, is that it "anticipates
regression to the mean". In other words, if we start from the prior that everyone is the same,
we will end up believing that the British people in the sample are less honest, and the Chinese 
people more honest, than the experimental data *per se* suggest.

@@

Table \@ref(tab:country-comparisons) shows Bayesian comparisons for selected countries. Even though
estimates have "shrunk" towards the pooled mean from the naive estimates, differences remain
significant, partly because the pooled prior has made posterior confidence intervals narrower.

@@

selected <- c("GB", "ZA", "GR", "BR", "CN")
ctry_comp <- ctry_comp[selected, selected]
cc_hux <- as_huxtable(ctry_comp, add_colnames = TRUE, add_rownames = TRUE) %>% 
      set_caption("(#tab:country-comparisons) Bayesian country comparisons. 
            Cells give the probability that row country has fewer liars than column country.") %>% 
      set_label("tab:country-comparisons") %>% 
      set_na_string("--") %>% 
      theme_basic()

@@

@@

Nevertheless, if I use GSV's method as stated, rather than their program, confidence intervals
remain too small, as shown in Table \@ref(tab:coverage-recalc).

@@

as_huxtable(coverage_recalc, add_colnames = TRUE) %>% 
      set_caption("(#tab:coverage-recalc) Proportion of true results within confidence interval, recalculated GSV method") %>% 
      set_label("tab:coverage-recalc") %>% 
      set_number_format(-1, everywhere, list(pct_format)) %>% 
      theme_article(header_col = FALSE)

@@

Figure \@ref(fig:coverage-pic) shows the proportions of true values within the confidence interval for the GSV method,
split by $N$, $P$ and confidence level. Dashed lines show the nominal confidence level.

Figure \@ref(fig:coverage-grp-pic) splits this still further, by the true proportion of liars within each
simulation. This makes the pattern clear: coverage gets worse as the proportion of liars increases (until 100% where it jumps back up as results become deterministic).

```{r coverage-pic, fig.cap = "GSV confidence interval coverage by confidence level, N and P"}

@@

```{r coverage-grp-pic, fig.cap = "GSV confidence interval coverage by confidence level, N, P and true proportion of liars"}

@@

Figure \@ref(fig:errors-by-true-prop) shows mean squared error by estimation method and 
true proportion of liars in the sample (True $L/N$), for $N$ &le; 50. The Bayesian method is best 
for all values of $L/N$ up to 80%.

```{r errors-by-true-prop, fig.cap = "Errors by method and true proportion of liars, N &le; 50"}
errors_true_prop <- sim_est %>%       
      filter(CI == 90, N <= 50) %>% 
      group_by(method, true_prop) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      rename(`True L/N` = true_prop, Method = method)

@@

```{r errors-disagg, include = FALSE}

@@

>>>>

---
title: "Lies and the lying liars who tell them"
subtitle: "Comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{setspace}
  - \usepackage{placeins}
  - \onehalfspacing
output:
  pdf_document:
    keep_tex: true
    md_extensions: +raw_attribute
    latex_engine: xelatex
  tufte::tufte_handout:
    latex_engine: xelatex
  bookdown::tufte_handout2:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
bibliography: bibliography.bib
link-citations: yes
---

@@

library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE)
options(huxtable.knitr_output_format = "latex")
source("bayesian-heads.R")
suppressPackageStartupMessages({
  library(Hmisc)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(broom)
  library(huxtable)
})

huxtable::set_default_properties(position = "left", top_padding = 7, bottom_padding = 12)
set.seed(19751027)

@@

I ran simulations to check the overall performance of the GSV confidence intervals. Simulations
parameters are shown in Table \ref{tab:params-table}. For each parameter combination, I ran 1000
simulations. "Liars" always reported heads. Non-liars reported tails with probability $P$ and
heads otherwise.

@@

huxtable(
        Parameter = c(
          "Sample size (N)", 
          "Proportion of liars (L/N)", 
          "Probability of bad random outcome (P)"
        ),
        Values    = c(
          paste(params$N, collapse = ", "),
          paste(params$Liars, collapse = ", "),
          paste(params$P, collapse = ", ")
        ),
        add_colnames = TRUE
      ) %>% 
      set_caption("Parameter values") %>% 
      set_label("tab:params-table") %>% 
      theme_article(header_col = FALSE)
```

For each simulation and confidence interval, I computed whether the confidence interval 
contained the true value of $L/N$. Table \ref{tab:coverage-overall} shows the results.

@@

cov_overall <- coverage %>% 
      ungroup() %>% 
      mutate(
        CI = paste0("CI ", CI, "%"),
        Method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian"),
        method = NULL
      ) %>% 
      spread(CI, mean_in_ci) %>% 
      arrange(factor(Method, levels = c("GSV", "Frequentist", "Bayesian")))

@@

ht_overall <- as_huxtable(cov_overall, add_colnames = TRUE) %>% 
      set_caption("Proportion of true results within confidence interval.") %>% 
      set_label("tab:coverage-overall") %>% 
      set_number_format(-1, -1, list(pct_format)) %>% 
      theme_article(header_col = FALSE)

@@

The problem is especially bad when there are many reports of heads. In this case there were probably
many true heads, so $T$ is high and the true sample size $N-T$ is low. Table \ref{tab:coverage-rn}
shows this. It splits the simulations by the proportion of reported heads, $R/N$. Coverage levels
fall off sharply as $R/N$ increases. Note that for fair coin flips, $R/N$ is typically greater than
0.5, in the simulations and in reality.

```{r coverage-rn}
coverage_rn <- sim_est %>% 
      filter(method=="gsv") %>% 
      group_by(
        `R/N:` = factor(Hmisc::cut2(R/N, cuts = c(0, 1/4, 1/2, 3/4, 1)))      
      ) %>% 
      mutate(pct_cases = n()/nrow(.)) %>% 
      group_by(`R/N:`, CI) %>% 
      summarize(
        mean_in_ci  = mean(within_ci, na.rm = TRUE), 
        `Percentage of cases` = pct_cases[1]
      ) %>% 
      mutate(CI = paste0("CI ", CI, "%")) %>% 
      spread(CI, mean_in_ci)

as_huxtable(coverage_rn, add_colnames = TRUE) %>% 
      set_caption("GSV confidence interval coverage by low and high R/N") %>% 
      set_label("tab:coverage-rn") %>% 
      set_escape_contents(2:3, 1, FALSE) %>% 
      set_number_format(-1, -1, list(pct_format)) %>% 
      theme_article(header_col = FALSE)
```

@@

Results are shown in Table \ref{tab:coverage-allmethods}. Both frequentist and  Bayesian methods
mostly achieve the nominal confidence level, with more than 90/95/99% of intervals containing the
true value of $Liars$. The exception is the frequentist 99% confidence interval, which is too
narrow.

```{r coverage-allmethods}
ht_overall %>% 
      set_caption("Coverage levels for GSV and alternative methods") %>% 
      set_label("tab:coverage-allmethods")
```

Frequentist confidence intervals could be less accurate when N is low, since that leads to more
sampling variation in the number of true heads. Table \ref{tab:coverage-N} checks this by looking
separately at simulations with $N=10$ and $N=50$. Frequentist 99% confidence intervals indeed appear
slightly too narrow for this range. Bayesian confidence intervals are fine. Both alternatives
perform much better than the GSV approach.

@@

as_huxtable(coverage_N, add_colnames = TRUE) %>% 
      set_rowspan(c(2, 4, 6), "Method", 2) %>% 
      set_valign(c(2, 4, 6), "Method", "top") %>% 
      set_number_format(-1, contains("CI"), list(pct_format)) %>% 
      set_caption("Confidence interval coverage by sample size") %>% 
      set_label("tab:coverage-N") %>% 
      theme_article(header_col = FALSE) %>% 
      set_bottom_border(6, 1, 0.4) # rowspan problems
```

@@

We can also compare the accuracy of point estimates of $\textit{Liars}$. Table
\ref{tab:point-estimates} shows the mean squared error for methods by different $N$. For low $N$,
the best method is Bayesian and the worst is Frequentist, with GSV in between. When $N$ gets large all
methods give about the same estimates and are equally accurate.

@@

sim_est %>% 
      filter(CI == 90) %>% 
      group_by(method, N) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("Mean squared errors by method and N") %>% 
      set_label("tab:point-estimates") %>% 
      theme_article(header_col = FALSE)
```

@@

```{r hj2016-pic, fig.cap = "Posterior means and confidence intervals for proportion dishonest in 15 countries. Crosses are naïve estimates.", fig.align="center"}

@@

Results are shown in Figure \ref{fig:hj2016-pic}, with means and 95% confidence intervals. There
is substantial shrinkage from the "naive" per-country estimates (using $2R/N - 1$). One of the strengths of empirical Bayes,  as @casella1985introduction points out, is that it "anticipates
regression to the mean". In other words, if we start from the prior that everyone is the same,
we will end up believing that the British people in the sample are less honest, and the Chinese 
people more honest, than the experimental data *per se* suggest.

@@

Table \ref{tab:country-comparisons} shows Bayesian comparisons for selected countries. Even though
estimates have "shrunk" towards the pooled mean from the naive estimates, differences remain
significant, partly because the pooled prior has made posterior confidence intervals narrower.

@@

selected <- c("GB", "ZA", "GR", "BR", "CN")
ctry_comp <- ctry_comp[selected, selected]
cc_hux <- as_huxtable(ctry_comp, add_colnames = TRUE, add_rownames = TRUE) %>% 
      set_caption("Bayesian country comparisons. 
            Cells give the probability that row country has fewer liars than column country.") %>% 
      set_label("tab:country-comparisons") %>% 
      set_na_string("--") %>% 
      theme_basic()

@@

\FloatBarrier

@@

Nevertheless, if I use GSV's method as stated, rather than their program, confidence intervals
remain too small, as shown in Table \ref{tab:coverage-recalc}.

@@

as_huxtable(coverage_recalc, add_colnames = TRUE) %>% 
      set_caption("Proportion of true results within confidence interval, recalculated GSV method") %>% 
      set_label("tab:coverage-recalc") %>% 
      set_number_format(-1, everywhere, list(pct_format)) %>% 
      theme_article(header_col = FALSE)

@@

Figure \ref{fig:coverage-pic} shows the proportions of true values within the confidence interval
for the GSV method, split by $N$, $P$ and confidence level. Dashed lines show the nominal confidence
level.

Figure \ref{fig:coverage-grp-pic} splits this still further, by the true proportion of liars within
each simulation. This makes the pattern clear: coverage gets worse as the proportion of liars
increases (until 100% where it jumps back up as results become deterministic).

```{r coverage-pic, fig.cap = "GSV confidence interval coverage by confidence level, N and P", fig.align="centre"}

@@

```{r coverage-grp-pic, fig.cap = "GSV confidence interval coverage by confidence level, N, P and true proportion of liars", fig.align="centre"}

@@

Figure \ref{fig:errors-by-true-prop} shows mean squared error by estimation method and 
true proportion of liars in the sample (True $L/N$), for $N$ of 10 and 50. The Bayesian method is best 
for all values of $L/N$ up to 80%.

```{r errors-by-true-prop, fig.cap = "Errors by method and true proportion of liars, N = 10 and 50", fig.align="centre"}
errors_true_prop <- sim_est %>%       
      filter(CI == 90, N <= 50) %>% 
      group_by(method, true_prop) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      ungroup() %>% 
      mutate(
        method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian")
      ) %>% 
      rename(`True L/N` = true_prop, Method = method)

@@

```{r errors-disagg, include = FALSE, fig.align="centre", fig.cap="Point estimate errors by R/N, N and P"}

@@

\FloatBarrier

# References
